Finding Quality Issues in SKOS Vocabularies

### 2012-03-23 analysis.tex
added numeric values in analysis tables

\todo{BH}{make clear that tool also provides reports not only numeric values}

@@

Our syntactical checks showed that only the AGROVOC vocabulary contained invalid datatypes and characters. After some minor syntax modifications, which don’t influence the evaluation against our criteria, we were also able to include this vocabulary. We also informed the AGROVOC team about this problem, which will try to fix it.

\todo{CM}{skip domain classification; this looks inconsistent and takes space.}

\begin{table}[h]
\caption{Analyzed vocabularies}
\begin{tabular}{lp{4.7cm}cccccc}
\textbf{Domain} & \textbf{Vocabulary} & \rotatebox{90}{\textbf{Concepts}} & \rotatebox{90}{\textbf{Lexical Rel.}} & \rotatebox{90}{\textbf{Semantic Rel.}} & \rotatebox{90}{\textbf{Aggregation Rel.}} & \rotatebox{90}{\textbf{Concept Schemes}} & \rotatebox{90}{\textbf{Collections}}\\
\toprule
Media & Gemeenschappelijke Thesaurus Audiovisuele Archieven (GTAA) & 171991 & 178776 & 50892 & 343980 & 9 & 0 \\
\hline
Science & Geonames Ontology & 671 & 671 & 0 & 671 & 9 & 0 \\
\hline
& Medical Subject Headings (MeSH) & 24626 & 150617 & 38858 & 0 & 0 & 0 \\
\hline
& Peroxisome Knowledge Base (PXV) & 2112 & 3628 & 2695 & 1716 & 1 & 0 \\
\hline
Government & Eurovoc & 6797 & 457788 & 18491 & 15512 & 128 & 0 \\
\hline
& Integrated Public Sector Vocabulary (IPSV) & 4732 & 7945 & 13843 & 4483 & 3 & 0 \\
\hline
Publications & Agrovoc & 31993 & 572138 & 63936 & 32043 & 1 & 0 \\
\hline
& DBpedia Categories & 743410 & 740352 & 1490316 & 0 & 0 & 0 \\
\hline
& University of Southampton Pressinfo & 1125 & 0 & 0 & 0 & 0 & 0 \\
\hline
& New York Times People (NYTP) & 4979 & 4979 & 0 & 4979 & 1 & 0 \\
\hline
& Library of Congress Subject Headings (LCSH) & 459182 & 746076 & 595754 & 815816 & 19 & 0 \\
\hline
Cross-domain & Meketre & 422 & 569 & 1698 & 6 & 2 & 0 \\
\hline
Economy & Thesaurus for Economics (STW) & 6524 & 31189 & 57907 & 6531 & 1 & 0 \\
\hline
& North American Industry Classification System (NAICS) & 4175 & 0 & 8684 & 2235 & 1 & 0 \\
\hline
Military & Thesaurus Landesverteidigungsakademie (LVAk) & 13411 & 17250 & 16346 & 0 & 0 & 0 \\
\bottomrule
\end{tabular}
\label{vocabs}
\end{table}

@@

\begin{table}[h]
\caption{Results of the proposed metrics}
\begin{tabular}{p{4cm}ccccccccccccccc}
\textbf{Metric} & \rotatebox{90}{\textbf{GTAA}} & \rotatebox{90}{\textbf{Geonames}} & \rotatebox{90}{\textbf{MeSH}} & \rotatebox{90}{\textbf{PXV}} & \rotatebox{90}{\textbf{Eurovoc}} & \rotatebox{90}{\textbf{IPSV}} & \rotatebox{90}{\textbf{Agrovoc}} & \rotatebox{90}{\textbf{DBpedia}} & \rotatebox{90}{\textbf{Pressinfo}} & \rotatebox{90}{\textbf{NYTP}} & \rotatebox{90}{\textbf{LCSH}} & \rotatebox{90}{\textbf{Meketre}} & \rotatebox{90}{\textbf{STW}} & \rotatebox{90}{\textbf{NAICS}} & \rotatebox{90}{\textbf{LVAk}} \\
\toprule
Omitted or Invalid Language Tags &&&&&&&&&&&&&&& \\

Incomplete Language Coverage &&&&&&&&&&&&&&& \\

Undocumented Concepts &&&&&&&&&&&&&&& \\

Label Conflicts &&&&&&&&&&&&&&& \\

@@

Loose Concepts & 162000 & 671 & 0 & 2 & 7 & 0 & 0 & 77062 & 1125 & 4979 & 172364 & 0 & 4 & 0 & 21 \\

@@

Link Target Unavailability &&&&&&&&&&&&&&& \\

@@

Undefined SKOS Property Usage &&&&&&&&&&&&&&& \\

>>>>

We implemented the qSKOS tool that can be used for computing the quality checking functions introduced in \ref{sec:criteria} and made it publicly available\footnote{\url{https://github.com/cmader/qSKOS/}}. All data in the Tables \ref{tab:vocs} and \ref{tab:results} have been obtained by utilizing this tool. However, qSKOS is not only capable of returning numeric values, but can also provide a detailed report listing all resources returned by the quality checking functions alongside with the potential cause of the issue.

@@

\begin{table}[h]
\label{tab:vocs}
\caption{Analyzed vocabularies}
\centering
\begin{tabular}{p{6cm}ccccccc}
\textbf{Vocabulary} & \rotatebox{90}{\textbf{Version/last mod.}} & \rotatebox{90}{\textbf{Concepts}} & \rotatebox{90}{\textbf{Labels}} & \rotatebox{90}{\textbf{Semantic Rel.}} & \rotatebox{90}{\textbf{Aggregation Rel.}} & \rotatebox{90}{\textbf{Concept Schemes}} & \rotatebox{90}{\textbf{Collections}}\\
\toprule
Gemeenschappelijke Thesaurus Audiovisuele Archieven (GTAA) & ? & 171991 & 178776 & 50892 & 343980 & 9 & 0 \\
\hline
Geonames Ontology & ? & 671 & 671 & 0 & 671 & 9 & 0 \\
\hline
Medical Subject Headings (MeSH) & ? & 24626 & 150617 & 38858 & 0 & 0 & 0 \\
\hline
Peroxisome Knowledge Base (PXV) & ? & 2112 & 3628 & 2695 & 1716 & 1 & 0 \\
\hline
Eurovoc & ? & 6797 & 457788 & 18491 & 15512 & 128 & 0 \\
\hline
Integrated Public Sector Vocabulary (IPSV) & ? & 4732 & 7945 & 13843 & 4483 & 3 & 0 \\
\hline
AGROVOC & 1.3 & 32035 & 620629 & 65934 & 32085 & 1 & 0 \\
\hline
DBpedia Categories & ? & 743410 & 740352 & 1490316 & 0 & 0 & 0 \\
\hline
University of Southampton Pressinfo & 2011/02/24 & 1125 & 0 & 0 & 0 & 0 & 0 \\
\hline
New York Times People (NYTP) & 2010/06/22 & 4979 & 4979 & 0 & 4979 & 1 & 0 \\
\hline
Library of Congress Subject Headings (LCSH) & ? & 459182 & 746076 & 595754 & 815816 & 19 & 0 \\
\hline
Meketre & 2011/07/07 & 422 & 569 & 1698 & 6 & 2 & 0 \\
\hline
Thesaurus for Economics (STW) & 8.06 & 6524 & 31189 & 57907 & 6531 & 1 & 0 \\
\hline
North American Industry Classification System (NAICS) & 2012 & 4175 & 0 & 8684 & 2235 & 1 & 0 \\
\hline
Thesaurus Landesverteidigungsakademie (LVAk) & n/a & 13411 & 17250 & 16346 & 0 & 0 & 0 \\
\bottomrule
\end{tabular}
\label{vocabs}
\end{table}

@@

\begin{table}[h]
\label{tab:results}
\caption{Results of the quality checking functions (cardinality of the result set)}
\begin{tabular}{p{4cm}ccccccccccccccc}
\textbf{Metric} & \rotatebox{90}{\textbf{GTAA}} & \rotatebox{90}{\textbf{Geonames}} & \rotatebox{90}{\textbf{MeSH}} & \rotatebox{90}{\textbf{PXV}} & \rotatebox{90}{\textbf{Eurovoc}} & \rotatebox{90}{\textbf{IPSV}} & \rotatebox{90}{\textbf{Agrovoc}} & \rotatebox{90}{\textbf{DBpedia}} & \rotatebox{90}{\textbf{Pressinfo}} & \rotatebox{90}{\textbf{NYTP}} & \rotatebox{90}{\textbf{LCSH}} & \rotatebox{90}{\textbf{Meketre}} & \rotatebox{90}{\textbf{STW}} & \rotatebox{90}{\textbf{NAICS}} & \rotatebox{90}{\textbf{LVAk}} \\
\toprule
Omitted or Invalid Language Tags &&&&&&& 0 & tba & 1224 & 0 & tba & 0 & 2 & 0 & 13411 \\

Incomplete Language Coverage &&&&&&& 32035 & tba & 0 & 0 & tba & 420 & 6456 & 0 & 0 \\

Undocumented Concepts &&&&&&& 32035 & tba & 1125 & 4094 & tba & 422 & 5236 & 3259 & 13411 \\

Label Conflicts &&&&&&& 2949 & tba & 0 & 0 & tba & 4 & 5 & 0 & 13 \\

@@

Orphan Concepts & 162000 & 671 & 0 & 2 & 7 & 0 & 0 & 77062 & 1125 & 4979 & 172364 & 0 & 4 & 0 & 21 \\

@@

Broken Links &&&&&&&&&&&&&&& \\

@@

Undefined SKOS Resources &&&&&&&&&&&&&&& \\

### 2012-03-23 criteria.tex
added numeric values in analysis tables

% Origin and rationale
SKOS defines a set of properties that link resources with RDF Literals, which are plain text natural language strings with an optional language tag. This includes the labeling properties \texttt{skos:prefLabel}, \texttt{skos:altLabel}, \texttt{skos:hiddenLabel} and also SKOS documentation properties, such as \texttt{skos:note} and subproperties thereof. Omitting language tags or using non-standardized, private language tags in a SKOS vocabulary could unintentionally limit the result set of language-dependent structural queries.
% Function defintion
We can check a SKOS vocabulary for omitted and invalid language tags by defining a quality checking function $oilt$ that iterates over all resources in $IR$ and finds those that have relations to plain literals with missing or invalid language tags, hence tags that are not defined in RFC3066\footnote{RFC 3066: Tags for the Identification of Languages \url{http://www.ietf.org/rfc/rfc3066.txt}}.

@@

% Origin and rationale
In the Linked Data context SKOS concepts should be linked with other related concepts on the Web, ``enabling seamless connections between data sets''\cite{Heath2011}. Similar to \emph{Missing In-Links}, this issue identifies the set of all conceptual resources that have no out-links. 
% Function definition
We provide the function \textit{mol} that iterates over all elements $c \in C$ and investigates \todo{CM}{mapping relations?} the host part of all related resource URIs. If none of these host names differ from the host part of $c$'s URI, it has no out-links and is thus included in the resources identified by this issue.

>>>>

% Origin and rationale
SKOS defines a set of properties that link resources with RDF Literals, which are plain text natural language strings with an optional language tag. This includes the labeling properties \texttt{skos:prefLabel}, \texttt{skos:altLabel}, \texttt{skos:hiddenLabel} and also SKOS documentation properties, such as \texttt{skos:note} and subproperties thereof. Omitting language tags or using non-standardized, private language tags in a SKOS vocabulary could unintentionally limit the result set of language-dependent structural queries.
% Function defintion
We can check a SKOS vocabulary for omitted and invalid language tags by defining a quality checking function $oilt$ that iterates over all resources in $IR$ and finds those that have relations to plain literals (using the above mentions SKOS properties) with missing or invalid language tags, hence tags that are not defined in RFC3066\footnote{RFC 3066: Tags for the Identification of Languages \url{http://www.ietf.org/rfc/rfc3066.txt}}.

@@

% Origin and rationale
In the Linked Data context SKOS concepts should be linked with other related concepts on the Web, ``enabling seamless connections between data sets''\cite{Heath2011}. Similar to \emph{Missing In-Links}, this issue identifies the set of all conceptual resources that have no out-links. 
% Function definition
We provide the function \textit{mol} that iterates over all elements $c \in C$ and investigates the host part of all related resource URIs. If none of these host names differ from the host part of $c$'s URI, it has no out-links and is thus included in the resources identified by this issue.
